\thetitle

### 2020-11-21 writing/results.tex
Editing results

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} on a --- specs
of fix. Each debugging scenario had a timeout limit of 4 minutes and an
out of memory limit of 6GB. The complete experiment required
approximately 270 hours of compute hours on our machine.

@@

Figure~\ref{subsec:experiment} shows that the answers for questions $Q_1$
to $Q_3$ are positive. In all three semantics, blame modes outperform
their corresponding exception mode (by 9.44\% to 12.7\%). The natural
exceptions mode is never more useful than natural blame and  transient
exceptions are more useful than transient first and transient last blame
in 2.19\% and 2.8\% of the scenarios respectively.

For the answers to $Q_*$, we compare the percentages in antisymmetric
positions in the matrix. Blame for all three semantics is significantly
more useful than Erasure exceptions (by 33\% to 35\%). Natural blame is
more useful than both versions of transient blame by a small percentage;
in 4.05\% of scenarios natural blame is more useful while in 1.67\% of the
scenarios transient blame is more useful. The transient first and
transient last blame are practically indistinguishable. Finally, there is
no clear winner between natural exceptions and transient exceptions
despite the theoretically advantageous additional checks of the natural
semantics.

@@

Figure~\ref{fig:effort-table} shows the distribution of programmer effort
for our sample of interesting debugging scenarios . Unlike the usefulness
comparisons above, these results do not generalize to the full population.
Still, as we discuss in section~\ref{subsec:effort}, they provide an
alternative view of the workings of the rational programmer.

There are
three immediate take-aways from the figure. First, the effort for successfully
debugging interesting scenarios (in green) for the random mode of the
rational program  follows  a normal distribution as expected.\footnote{The
random mode distribution is indistinguishable for all three semantics and thus the figure~\ref{fig:effort-table} 
only shows one plot.} In contrast, in the other modes, successful effort coalesces at
the left side of the plot, meaning that in most cases the programmer needs
to type a single component to debug a scenario. Even in the minority of
cases where a blame or exception mode has to work harder, the number of
modules that that need to be typed is small relative to the random mode.

A second point of interest is that the exceptions of
the erasure semantics either help the rational programmer immediately or 
the rational programmer fails to debug a scenario altogether (in red). This shouldn't
come as a surprise; no matter how many type annotations the programmer
adds to an erasure program, if the type checker doesn't reject the
program, running the program produces always the same outcome. Thus an
exception from the erasure's runtime has to point to the buggy module with
the first try. Otherwise the rational programmer types an irrelevant
module, runs the program again and the exception points again to the
already typed module.

A third notable piece of information from
the figure is that every mode has failed debugging scenarios, not just
erasure. This should
not come as a surprise to the astute
reader. After all, as we
discuss above, some modes are more successful than others in certain
scenarios. Indeed any mode can fail if running a scenario results in an
exception and the exception carries no useful information about which
module the programmer should type next, e.g., because the stack trace 
does not contain frames from any module of the program. This situation
can even happen immediately, which explains failures with 0 effort. 
However  while the majority of
failures to debug failed scenarios follow this pattern, a few do not. The next section takes a
closer look at these unexpected failures of the rational programmer as part of the lessons we
extract from our experimental results.

>>>>

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} using a machine with two Intel Xeon Gold 6258R processors and 500GB of memory.
Each debugging scenario had a timeout limit of 4 minutes and a
memory limit of 6GB. The complete experiment required
approximately 270 hours of compute hours on our machine.

@@

Figure~\ref{subsec:experiment} shows that the answers for questions $Q_1$
to $Q_3$ are positive. In all three semantics, blame modes outperform
their corresponding exception mode (by 9.44\% to 12.7\%). The Natural
exceptions mode is never more useful than Natural blame, and  Transient
exceptions are more useful than Transient first and Transient last blame
in 2.19\% and 2.8\% of the scenarios respectively.

For the answers to $Q_*$, we compare the percentages in antisymmetric
positions in the matrix. Blame for all three semantics is significantly
more useful than Erasure exceptions (by 33\% to 35\%). Natural blame is
more useful than both versions of Transient blame by a small percentage;
in 4.05\% of scenarios Natural blame is more useful while in 1.67\% of the
scenarios Transient blame is more useful. The Transient first and
Transient last blame are practically indistinguishable. Finally, there is
no clear winner between Natural exceptions and Transient exceptions
despite the theoretically advantageous additional checks of the Natural
semantics.

@@

Figure~\ref{fig:effort-table} shows the distribution of programmer effort
for our sample of interesting debugging scenarios. Unlike the usefulness
comparisons above, these proportions do not generalize to a representation of the full population.
Still, as we discuss in section~\ref{subsec:effort}, they provide an
alternative view of the workings of the rational programmer.

There are
three immediate take-aways from the figure. First, the effort for successfully
debugging interesting scenarios (in green) for the random mode of the
rational programmer follows  a normal distribution, as expected.\footnote{The
random mode distribution is indistinguishable for all three semantics and thus the figure~\ref{fig:effort-table} 
only shows one plot.} In contrast, in the other modes, successful effort coalesces at
the left side of the plot, meaning that in most cases the programmer needs
to type a single component to debug a scenario. Even in the minority of
cases where a blame or exception mode has to work harder, the number of
modules that that need to be typed is small relative to the random mode.

A second point of interest is that the exceptions of
the Erasure semantics either help the rational programmer immediately or 
the rational programmer fails to debug a scenario altogether (in red). This shouldn't
come as a surprise; no matter how many type annotations the programmer
adds to an Erasure program, if the type checker doesn't reject the
program, running the program always produces the same outcome. Thus an
exception from the runtime has to point to the buggy module with
the first try. Otherwise the rational programmer types an irrelevant
module, runs the program again and the exception points again to the
already typed module.

A third notable piece of information from
the figure is that every mode has failed debugging scenarios, not just
Erasure. This should
not come as a surprise to the astute
reader. After all, as we
discuss above, some modes are more successful than others in certain
scenarios. Indeed any mode can fail if running a scenario results in an
exception and the exception carries no useful information about which
module the programmer should type next, e.g., because the stack trace 
does not contain frames from any module of the program, or only typed modules. This situation
can even happen immediately, which explains failures with 0 effort. 
While the majority of
failures to debug failed scenarios follow this pattern, a few do not. The next section takes a
closer look at these unexpected failures of the rational programmer as part of the lessons we
extract from our experimental results.
